% This line sets the project root file.
% !TEX root = Notes_Gauging_Defects.tex
% !TEX spellcheck = en_US

Many important discoveries in condensed matter physics during the past decades have arisen from the study of \emph{topological states of matter} (see, for example, \cite{Wen07,NSSFD08,HK10,QZ11}). For a long time it was thought that all continuous phase transitions could be described by the Ginzburg-Landau theory of symmetry breaking. In the late 1980s, however, physicists discovered that some systems can exhibit a new kind of order going beyond the usual symmetry breaking paradigm. This new kind of order -- \emph{topological order} \cite{Wen90} -- soon became an interesting topic in its own right, useful for the description of different quantum Hall states \cite{WN90}. More recently, topologically ordered systems gave rise to \emph{topological quantum computation}, because of their remarkable properties: these locally interacting systems exhibit emergent global properties protected against environmental noise. This makes them a promising platform for the robust encoding, storage, and manipulation of quantum systems \cite{DKLP2002,Kit03,NSSFD08,Ter15,PY15,BLPSW16}.

However, in many phases, which includes those most suitable for experimental realization, the quantum computational power of a topologically ordered system is severely limited. To ameliorate this, defects in topologically ordered systems were considered \cite{RH07,Bombin2010,KK12,FSV13,BJQ13b,BASP14,JPSV15,DIP16,CCW16,BBD17,CCW17b,CCW17,BLKW17,KPEB18,ET19}, since a theory which includes defects can describe topological phases with enhanced computational power \cite{Freedman1998,FLW02b,FLW02,FKLW02}. For example, it is possible to use topologically ordered systems which are not universal for quantum computation to realize a computationally universal braiding gate set by adding defects \cite{BJQ13}.

In condensed matter physics, there is another approach to adding defects, which even in the case of simple and well-understood quantum systems substantially enriches their physics. Consider a simple example of fermions moving freely in one spatial dimension: the addition of a single defect -- an \emph{impurity} -- leads to a substantially richer and more complex system exhibiting Kondo-type effects \cite{andersonLocalizedMagneticStates1961,hewsonKondoProblemHeavy1997}. The study of such impurity models by Wilson \cite{wilsonRenormalizationGroupCritical1975} directly led to the revolutionary development of the density matrix renormalization group (DMRG) \cite{whiteDensityMatrixFormulation1992}, and the subsequent tensor network revolution \cite{bridgemanHandwavingInterpretiveDance2017}.

In the presence of additional symmetries, the phases of quantum systems have an even finer classification \cite{Wen2002,SRFL08,Kitaev2009,FK10,CGW11,FK11,TPB11,LS12,LV12,FM13,EH13,NCMT14,WPS14,K14,F14,EN14,MFCV15,BRSX15,LV16}. More precisely, it is possible for two quantum states that are equivalent when there is no additional symmetry present to be distinct in the presence of additional symmetries. This phenomenon is referred to as \emph{symmetry-protected topological} order (SPT) \cite{CGLW13,Yoshida2015,Yoshida2017} if the gapped phase is trivial in the absence of any symmetry, or as \emph{symmetry-enriched topological} (SET) order \cite{ENO10,MR13,WBV17} if the phase is topologically nontrivial. As far as defects are concerned, not only the distinct phases of matter acquire a finer classification in the presence of additional symmetries, but also the class of possible defects becomes larger. 

It is possible to to transform a topologically ordered system in the presence of a global symmetry to a system with a local gauge invariance. This is referred to as ``gauging the symmetry''. This is convenient for several reasons. For example, many properties of the ungauged system can be understood by looking at the properties of the gauged theory \cite{LG12,HW12,Swingle2014,CG14}. Furthermore, it may give insights into the quantum phase transitions between two different topological phases of matter \cite{BS09,BW10,BW11,BSS12}. 
The study of the fusion properties of defects in general quantum phases -- particularly when they are allowed to be mobile -- is deeply connected with global and local symmetries and the formalism of \emph{G-crossed braided tensor categories} \cite{BBCW14}. This connection has been studied from a mathematical point of view in various works \cite{T00,ENO10,Turaev2010,CGPW16,EMJP18,CSZW18,D19,BJ19}. The study of models involving mobile defects -- analogues of impurity models for anyons -- with nontrivial dynamics lies at the very cutting edge of current research. Such models promise to provide new phases of matter even in $(1+1)$ dimensions, going beyond the SPT classification.

From a physical perspective, one can perform a microscopic approach in order to make sense of a dynamical theory of defects for quantum (spin) systems. We restrict our attention to defects modeled by absent/missing quantum spins. (Extending this idea to more general defects modeled by different Hilbert spaces then becomes a straightforward task.)

We build up to a theory of dynamical defects in stages. In the first stage we focus on the problem of modeling an \emph{indefinite} number of distinguishable spins. Building on this we can propose a kinematical space to model arbitrary numbers of quantum spins at various locations. This allows us to then consider the case of missing-spin defects for topologically ordered systems.

\subsection{Describing an indefinite number of distinguishable quantum spins}

When we study quantum theory we learn that to describe the Hilbert space $\mathcal{H}_{AB}$ of systems comprised of two or more distinguishable subsystems we should use the tensor product:
\begin{equation}
\mathcal{H}_{AB} \cong \mathcal{H}_A\otimes \mathcal{H}_B,
\end{equation}
where $\mathcal{H}_A$ and $\mathcal{H}_B$ are the Hilbert spaces for subsystems $A$ and $B$, respectively. We also learn that to describe a system with an \emph{indefinite} number of \emph{indistinguishable} particles we should use \emph{Fock space}. The construction of Fock space is often presented in a confusing way which intermixes the roles of exchange symmetry and indefinite particle number.

What is not so commonly taught, but should be, is the arguably conceptually simpler construction of \emph{distinguishable Fock space} (see \cite{OsborneVideoLectureAQT} for a course where this is explained), appropriate for describing a collection of an indefinite number of distinguishable particles. Subsequently, one can obtain Bose-/Fermi-Fock space by imposing an equivalence relation under particle exchange. This approach emphasizes the separation of concerns, and, as we will see, also affords us considerable freedom in modeling more exotic situations such as the description of dynamical defects.

Suppose we want to model the following situation. Imagine we have a system comprised of $N$ distinguishable quantum spins with local Hilbert space $\mathbb{C}^d$. As per the composite-system axiom of quantum mechanics, we model such a system with the Hilbert space
\begin{equation}
\mathcal{H}_N \cong \bigotimes_{j=1}^N \mathbb{C}^d.
\end{equation}
This Hilbert space describes all states of the $N$ \emph{individually identifiable and addressable} quantum spins. The $j$th tensor factor corresponds to the internal states of the $j$th spin.

What if we don't know beforehand how many particles we'll end up with? Let's assume that, even though we don't know how many particles we'll have, we will still be able to individually identify and address the particles \footnote{If you want a physical example for such a system: imagine a $1D$ optical lattice in which either zero or one two-level atoms can be present at each lattice site. Further, assume that the atoms line up in a contiguous line, with no gaps, i.e., we impose some additional potential gradient. Although we don't know how many particles we'll get, we can still identify and address the particles via lattice location.}. How should we incorporate the additional quantum number $N$, describing the size of the collection? The answer is via the \emph{direct sum}: we take the direct sum over $N$ of the Hilbert space $\mathcal{H}_N$ for $N$ distinguishable particles. This results in \emph{distinguishable Fock space}
\begin{equation}
\mathfrak{F}(\mathbb{C}^d) \cong \bigoplus_{N=0}^\infty \mathcal{H}_N\cong \bigoplus_{N=0}^\infty (\underbrace{\mathbb{C}^d\otimes \mathbb{C}^d\otimes \cdots \otimes \mathbb{C}^d}_{\text{$N$ factors}}).
\end{equation}
By convention and assumption the space describing zero particles, the \emph{vacuum}, is assigned the Hilbert space $(\mathbb{C}^d)^{\otimes 0} \cong \mathbb{C}$.

It is now easy to incorporate additional constraints on the numbers of particles, e.g., to describe a system comprised of \emph{either} zero distinguishable quantum spins \emph{or} one distinguishable quantum spin we would use
\begin{equation}
\mathfrak{F}_{\le 1} (\mathbb{C}^d) \cong \mathbb{C}\oplus \mathbb{C}^d.
\end{equation}
We could call this the Hilbert space of a ``maybe'' quantum spin.

Let's now consider the central situation for this work. How should we describe a quantum lattice of $n$ sites, where either a single quantum spin is present at a site, or not? (We call the case where a spin is absent a \emph{defect}.) According to the discussion above we should simply tensor up $n$ ``maybe'' quantum spins:
\begin{equation}
\mathfrak{F}_{\le n}(\mathbb{C}^d) \equiv \bigotimes_{j=0}^n \mathfrak{F}_{\le 1}(\mathbb{C}^d) \cong (\mathbb{C}\oplus \mathbb{C}^d)^{\otimes j}.
\end{equation}
Expanding out the tensor factors leads to the equivalent definition
\begin{equation}
\mathfrak{F}_{\le n}(\mathbb{C}^d) \equiv \bigoplus_{j=0}^n \mathbb{C}^{\binom{n}{j}}\otimes \mathcal{H}_j.
\end{equation}
A quick way to convince yourself of the validity of this second representation just take the dimension of both definitions. In the first case we get $\dim(\mathfrak{F}_{\le n}(\mathbb{C}^d)) = (d+1)^n$ and in the second we find
\begin{equation}
\dim(\mathfrak{F}_{\le n}(\mathbb{C}^d)) = \sum_{j=0}^n \binom{n}{j} d^j,
\end{equation}
which is identical due to the binomial theorem. We call the space arising in the second representation
\begin{equation}
\mathbb{C}^{\binom{n}{j}} \cong \mathcal{K}_{j};
\end{equation}
it may be interpreted as the configuration space of $j$ hard-core identical scalar particles arranged on a system of $n$ possible locations. Our total space still corresponds to \emph{distinguishable} particles; the tensor factor $\mathcal{K}_j$ takes care of identifying which locations are occupied with spins and the second tensor factor represents the actual distinguishable spins at those locations.

The Hilbert space $\mathfrak{F}_{\le n}(\mathbb{C}^d)$ supplies us with just the kinematical data to describe a system of distinguishable particles. To incorporate additional dynamical information we must specify additional \emph{dynamical} data.

\subsection{Modeling dynamically evolving defects}
\label{Intro_dynamics}

We specify dynamics in quantum mechanics most directly via a one-parameter family of unitary operators $U_t:\mathcal{H}\rightarrow\mathcal{H}$; usually the family is continuous enough that we can represent them via a \emph{Hamiltonian} $H$:
\begin{equation}
U_t = e^{-it H}.
\end{equation}
To introduce dynamics for our defect system we hence just need a hermitian matrix acting on $\mathfrak{F}_{\le n}(\mathbb{C}^d)$. This is easy enough in the abstract, however, particularly for topologically ordered systems we introduce dynamical information more indirectly.

In the context of topologically ordered models such as Kitaev's toric code \cite{Kit03} we typically introduce dynamical information indirectly by describing the system via the ground eigenspace $\mathcal{V}\subset \mathcal{H}$ of a specific Hamiltonian $H$. In the case of the toric code for $N\times N$ quantum spins arranged on the torus, the ground eigenspace $\mathcal{V}_{\mathbb{T}}$ of the toric code Hamiltonian is four dimensional.

Continuing in the context of the toric code, let's suppose we now have a lattice with a missing-qubit defect at some edge $e$ \cite{BLKW17}, see Fig. \ref{fig:1defect}. One can define a toric-code Hamiltonian for this new punctured torus; restricting to its ground eigenspace yields a \emph{four-dimensional subspace} $\mathcal{V}_{\mathbb{T}\setminus e}$. There is no obstruction to describing the ground eigenspace for two, three, etc.\ missing-qubit defects. What results is a rather intricate combinatorial problem, as, depending on where the defects are located relative to each other, one gets higher or lower dimensional ground eigenspace. To see an example of the intricacies that easily result, consider the case of two missing edges: these can either be adjacent to each other or distant. In the latter case, depicted in Fig. \ref{fig:2distdefects}, the ground space for this system is $(\mathbb{C}^2)^{\otimes 3}$ while in the case of adjacent defects we have a single larger puncture (with smooth boundary), see Fig. \ref{fig:2adjdefects}, and the ground eigenspace is only $(\mathbb{C}^2)^{\otimes 2}$.

\begin{figure}
	\centering
	\begin{subfigure}[t]{0.3\linewidth}
		\centering
		\begin{tikzpicture}[scale=0.75,baseline=(current bounding box.center)]
		%horizontal lines
		\draw (-0.25,0) -- (3.25,0);
		\draw (-0.25,-1) -- (3.25,-1);
		\draw (-0.25,-2) -- (1,-2);
		\draw (2,-2) -- (3.25,-2);
		\draw (-0.25,-3) -- (3.25,-3);
		%vertical lines
		\draw (0,-3.25) -- (0,0.25);
		\draw (1,-3.25) -- (1,0.25);
		\draw (2,-3.25) -- (2,0.25);
		\draw (3,-3.25) -- (3,0.25);
		\end{tikzpicture}
		\caption{One missing qubit defect on a toric code lattice.}\label{fig:1defect}
	\end{subfigure}%\vspace{20pt}	
\hspace{0.2\linewidth}
	\begin{subfigure}[t]{0.3\linewidth}
		\centering
		\begin{tikzpicture}[scale=0.75,baseline=(current bounding box.center)]
		%horizontal lines
		\draw (-0.25,0) -- (3.25,0);
		\draw (-0.25,-1) -- (1,-1);
		\draw (2,-1) -- (3.25,-1);
		\draw (-0.25,-2) -- (1,-2);
		\draw (2,-2) -- (3.25,-2);
		\draw (-0.25,-3) -- (3.25,-3);
		%vertical lines
		\draw (0,-3.25) -- (0,0.25);
		\draw (1,-3.25) -- (1,0.25);
		\draw (2,-3.25) -- (2,0.25);
		\draw (3,-3.25) -- (3,0.25);
		\end{tikzpicture}
		\caption{Two adjacent missing qubit defects on a toric code lattice.}\label{fig:2adjdefects}
	\end{subfigure}\vspace{20pt}
	
	\begin{subfigure}[t]{0.3\linewidth}
		\centering
		\begin{tikzpicture}[scale=0.75,baseline=(current bounding box.center)]
		%horizontal lines
		\draw (-0.25,0) -- (3.25,0);
		\draw (-0.25,-1) -- (2,-1);
		\draw (3,-1) -- (3.25,-1);
		\draw (-0.25,-2) -- (1,-2);
		\draw (2,-2) -- (3.25,-2);
		\draw (-0.25,-3) -- (3.25,-3);
		%vertical lines
		\draw (0,-3.25) -- (0,0.25);
		\draw (1,-3.25) -- (1,0.25);
		\draw (2,-3.25) -- (2,0.25);
		\draw (3,-3.25) -- (3,0.25);
		\end{tikzpicture}
		\caption{Two distant missing qubit defects on a toric code lattice.}\label{fig:2distdefects}
	\end{subfigure}
	\caption{Depiction of different situations where missing qubit defects are inserted on a toric code lattice. In the case of more than one missing link the ground space for the system depends on whether the defects are adjacent to each other.}
\end{figure}

%\begin{figure}
%	\centering
%	\stackunder[5pt]{
%	\begin{tikzpicture}[scale=0.75,baseline=(current bounding box.center)]
%	%horizontal lines
%	\draw (-0.25,0) -- (3.25,0);
%	\draw (-0.25,-1) -- (3.25,-1);
%	\draw (-0.25,-2) -- (1,-2);
%	\draw (2,-2) -- (3.25,-2);
%	\draw (-0.25,-3) -- (3.25,-3);
%	%vertical lines
%	\draw (0,-3.25) -- (0,0.25);
%	\draw (1,-3.25) -- (1,0.25);
%	\draw (2,-3.25) -- (2,0.25);
%	\draw (3,-3.25) -- (3,0.25);
%	\end{tikzpicture}
%		}
%		{(a) One missing qubit defect on a toric code lattice.}\label{fig:1defect}
%	\hspace{40pt}
%	\stackunder[5pt]{
%		\begin{tikzpicture}[scale=0.75,baseline=(current bounding box.center)]
%		%horizontal lines
%		\draw (-0.25,0) -- (3.25,0);
%		\draw (-0.25,-1) -- (1,-1);
%		\draw (2,-1) -- (3.25,-1);
%		\draw (-0.25,-2) -- (1,-2);
%		\draw (2,-2) -- (3.25,-2);
%		\draw (-0.25,-3) -- (3.25,-3);
%		%vertical lines
%		\draw (0,-3.25) -- (0,0.25);
%		\draw (1,-3.25) -- (1,0.25);
%		\draw (2,-3.25) -- (2,0.25);
%		\draw (3,-3.25) -- (3,0.25);
%		\end{tikzpicture}
%	}
%	{(b) Two adjacent missing qubit defects on a toric code lattice.}\label{fig:2adjdefects}
%	\hspace{40pt}
%	\stackunder[5pt]{
%		\begin{tikzpicture}[scale=0.75,baseline=(current bounding box.center)]
%		%horizontal lines
%		\draw (-0.25,0) -- (3.25,0);
%		\draw (-0.25,-1) -- (2,-1);
%		\draw (3,-1) -- (3.25,-1);
%		\draw (-0.25,-2) -- (1,-2);
%		\draw (2,-2) -- (3.25,-2);
%		\draw (-0.25,-3) -- (3.25,-3);
%		%vertical lines
%		\draw (0,-3.25) -- (0,0.25);
%		\draw (1,-3.25) -- (1,0.25);
%		\draw (2,-3.25) -- (2,0.25);
%		\draw (3,-3.25) -- (3,0.25);
%		\end{tikzpicture}
%	}
%	{(c) Two distant missing qubit defects on a toric code lattice.}\label{fig:2distdefects}
%	\caption{Depiction of different situations where missing qubit defects are inserted on a toric code lattice. In the case of more than one missing link the ground space for the system depends on whether the defects are adjacent to each other.}
%\end{figure}

We can incorporate such \emph{indirect} dynamical information into our kinematical space $\mathcal{F}_{\le n}(\mathbb{C}^d)$ as follows. We start by writing out 
\begin{equation}
\mathcal{V}_{\le n}(\mathbb{C}^d) = \bigoplus_{e_1,e_2, \ldots, e_{N^2} =0}^1 \mathcal{V}_{e_1,e_2,\ldots, e_{N^2}},
\end{equation}
where $\mathcal{V}_{e_1,e_2,\ldots, e_{N^2}}$ is the ground eigenspace for the toric code with a missing-edge defect at every edge $e_j$ with $e_j=0$. Writing out the first terms of this big direct sum gives us
\begin{equation}
\mathcal{V}_{\le n}(\mathbb{C}^d) = (\mathbb{C}^2\otimes \mathbb{C}^2)\oplus \mathbb{C}^{N^2}\otimes (\mathbb{C}^2\otimes \mathbb{C}^2\otimes \mathbb{C}^2)\oplus \cdots.
\end{equation}
The first direct summand corresponds to the ground eigenspace of the toric code without defects, the second summand corresponds to the $N^2$ possible single defects, etc.

The full direct-sum structure of $\mathcal{V}_{\le n}(\mathbb{C}^d)$ is not easy to describe. Indeed, this is why we use a category theoretic language to discuss these problems. 

So far, the space $\mathcal{V}_{\le n}(\mathbb{C}^d)$ is the low-energy configuration space for a topologically ordered system, the toric code, on a lattice with an indefinite number of missing-edge defects. In this case, we assume we somehow \emph{can know} where the defects are located. However, for a full dynamical theory, this assumption is potentially unjustified.

To incorporate the effects of such limited detection ability we impose an \emph{equivalence relation} on $\mathcal{V}_{\le n}(\mathbb{C}^d)$ where we identify ``physically indistinguishable'' configurations of defects. There are many possible notions of indistinguishability: this is an \emph{operational} notion and must be justified on a case by case basis. 

One particular, rather coarse, notion of indistinguishability may be justified as follows. Suppose we have a state $|\phi\rangle$ of the system for a defect configuration $e_1,e_2, \ldots, e_{N^2}$ and another $|\psi\rangle$ for a defect configuration with the same number of defects at possible different locations $e_1',e_2',\ldots, e'_{N^2}$. We say that the two states $|\phi\rangle$ and $|\psi\rangle$ are \emph{equivalent} if there is a unitary circuit which can transform $|\phi\rangle$ to $|\psi\rangle$. This, in particular, requires that the subspaces $\mathcal{V}_{e_1,e_2,\ldots,e_{N^2}}$ and $\mathcal{V}_{e_1',e_2',\ldots,e_{N^2}'}$ which contain $|\phi\rangle$ and $|\psi\rangle$ have the same dimensions.

This equivalence relation collapses many of the direct sums appearing in $\mathcal{V}_{e_1,e_2,\ldots,e_{N^2}}$, e.g., after applying the equivalence relation the first two summands above become
\begin{equation}
\left(\mathcal{V}_{\le n}(\mathbb{C}^d)/\sim\right) = (\mathbb{C}^2\otimes \mathbb{C}^2)\oplus (\mathbb{C}^2\otimes \mathbb{C}^2\otimes \mathbb{C}^2)\oplus \cdots.
\end{equation}

Of course, the above description is not only restricted to the toric code case. However, even in this simple example writing out the whole ground eigenspace and applying the equivalence relation to it results in a highly intricately combinatorial problem. That is why we exploit concepts from category theory to model defects in the system. By this approach we are able to handle more complex systems efficiently. In fact, it enables us to study defects in arbitrary quantum systems, whenever the underlying fusion algebra has the structure of a unitary fusion category. In this work, we will focus on the example of a one-dimensional quantum spin chain with $\Vec(\Z/2\Z)$ fusion rules. We explicitly show, how the above construction works and how ideas and techniques from category theory can be used to model dynamical defects in such a one-dimensional spin system. Here, defects are modeled by an invertible bimodule of the underlying category, see Subsec.~\ref{subsec:VecZ2}. Using generalized tube algebra techniques \cite{ocneanu}, we construct explicit trivalent vertices for the gauged theory. Consistent with the literature \cite{TY,ENO10,Bombin2010,BBCW14,WBV17}, we find that the $F$-symbols of the gauged theory for the $\Vec(\Z/2\Z)$ spin chain (computed using these explicit vertices) coincide exactly with those of the Ising category. To the best of our knowledge, this is the first time such a result has been computed directly using the tube algebra. Furthermore, defining a golden-chain-like Hamiltonian \cite{Feiguin2007} for this spin chain results in the transverse Ising model.

%In our work, the goal is to make physical and microscopic sense of a dynamical theory of defects for a one-dimensional quantum spin chain with $\Vec(\Z/2\Z)$ fusion rules. While we are focusing on this particular case study, the framework we establish can easily be extended to arbitrary quantum spin chains. Define the kinematical data for the system with an indefinite number of defects is straightforward via the well-known Fock space construction, whereas imposing dynamics is more complicated. In the context of topologically ordered systems, such as the considered $\Vec(\Z/2\Z)$ spin chain, dynamical information is typically introduced via the ground eigenspace of a specific Hamiltonian that represents the topological properties of the system. A common example is Kitaev's toric code for $N\times N$ quantum spins arranged on the torus, which exhibits a four-dimensional ground eigenspace. Defects in this system could be modeled by absent quantum spins, resulting in ground eigenspaces of differing dimensions. When considering more than one defect (i.e., missing spins) one also has to distinguish between cases where the defects are next to each other and those where they are spatially separated. This can result in a complicated combinatorial problem as shown in Subsec.~\ref{Kin_Dyn}. Therefore, we use tools from category theory to simplify explicit computations.

This paper is structured as follows: In Sec.~\ref{S:defs}, we provide definitions and background required for the remainder of the paper. In particular, we need the concepts from generalized tube algebras provided here to be able to study explicit examples of models with dynamical defects. Afterwards, we transfer the previously introduced concepts to a spin chain with $\Vec(\Z/2\Z)$ fusion rules in Sec.~\ref{Ising}, which enables us to compute its dynamical data. This leads to a realization of the transverse Ising model. Finally, we conclude in Sec.~\ref{Conclusion}.